Do we need detailed demographic data to forecast plant population responses to climate change?

### 2016-03-12 manuscript/manuscript.Rmd
updates ms with JECOL style abstract

Summary
-------
1. Rapid climate change has generated growing interest in forecasts of future population trajectories. Traditional population models, typically built using detailed demographic observations from one study site, can address climate change impacts at one location, but are difficult to scale up to the landscape and regional scales relevant to management decisions. An alternative is to build models using population-level data that is much easier to collect over broad spatial scales than individual-level data. However, such models ignore the fact that climate drives population growth through its influence on individual performance. 
2. We test the consequences of aggregating individual responses when forecasting climate change impacts on two perennial grass species (*Pascopyrum smithii* and *Sprobolis flexuosus*). We parameterized two population models for each species, one based on individual-level data (survival, growth and recruitment) and one on population-level data (percent cover), and compared their accuracy, precision, and sensitivity to climate variables. For both models we used Bayesian ridge regression to identify the optimal predictive model in terms of climate covariate strengths. 
3. The individual-level model was more accurate and precise than the aggregated model when predicting out-of-sample observations. When comparing climate effects from both models, the population-level model missed important climate effects from at least one vital rate for each species. Increasing the sample size at the population-level would not necessarily reduce forecast uncertainty; the way to reduce uncertainty is to capture unique climate dependence of individual vital rates. 
4. Our analysis indicates that there is no shortcut to forecasting climate change impacts on plant populations — detailed demographic data is essential. Despite the superiority of the individual-level model, the forecasts it generated still were too uncertain to be useful for decision-makers. We need new methods to collect demographic data efficiently across environmental gradients in space and time.

>>>>

Summary
-------
1. Rapid climate change has generated growing interest in forecasts of future population trajectories. Traditional population models, typically built using detailed demographic observations from one study site, can address climate change impacts at one location, but are difficult to scale up to the landscape and regional scales relevant to management decisions. 
2. An alternative is to build models using population-level data that is much easier to collect over broad spatial scales than individual-level data. However, such models ignore the fact that climate drives population growth through its influence on individual performance. 
3. We test the consequences of aggregating individual responses when forecasting climate change impacts on four perennial grass species. We parameterized two population models for each species, one based on individual-level data (survival, growth and recruitment) and one on population-level data (percent cover), and compared their accuracy, precision, and sensitivity to climate variables. For both model types we used Bayesian ridge regression to identify the optimal predictive model in terms of climate covariate strengths. 
4. The individual-level model was more accurate and precise than the aggregated model when predicting out-of-sample observations. When comparing climate effects from both models, the population-level model missed important climate effects from at least one vital rate for each species. Increasing the sample size at the population-level would not necessarily reduce forecast uncertainty; the way to reduce uncertainty is to capture unique climate dependence of individual vital rates. 
5. *Synthesis* Our analysis indicates that there is no shortcut to forecasting climate change impacts on plant populations — detailed demographic data is essential. Despite the superiority of the individual-level model, the forecasts it generated still were too uncertain to be useful for decision-makers. We need new methods to collect demographic data efficiently across environmental gradients in space and time.

### 2016-03-12 manuscript/manuscript.tex
updates ms with JECOL style abstract

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Rapid climate change has generated growing interest in forecasts of
  future population trajectories. Traditional population models,
  typically built using detailed demographic observations from one study
  site, can address climate change impacts at one location, but are
  difficult to scale up to the landscape and regional scales relevant to
  management decisions. An alternative is to build models using
  population-level data that is much easier to collect over broad
  spatial scales than individual-level data. However, such models ignore
  the fact that climate drives population growth through its influence
  on individual performance.
\item
  We test the consequences of aggregating individual responses when
  forecasting climate change impacts on two perennial grass species
  (\emph{Pascopyrum smithii} and \emph{Sprobolis flexuosus}). We
  parameterized two population models for each species, one based on
  individual-level data (survival, growth and recruitment) and one on
  population-level data (percent cover), and compared their accuracy,
  precision, and sensitivity to climate variables. For both models we
  used Bayesian ridge regression to identify the optimal predictive
  model in terms of climate covariate strengths.
\item
  The individual-level model was more accurate and precise than the
  aggregated model when predicting out-of-sample observations. When
  comparing climate effects from both models, the population-level model
  missed important climate effects from at least one vital rate for each
  species. Increasing the sample size at the population-level would not
  necessarily reduce forecast uncertainty; the way to reduce uncertainty
  is to capture unique climate dependence of individual vital rates.
\item
  Our analysis indicates that there is no shortcut to forecasting
  climate change impacts on plant populations --- detailed demographic
  data is essential. Despite the superiority of the individual-level
  model, the forecasts it generated still were too uncertain to be
  useful for decision-makers. We need new methods to collect demographic
  data efficiently across environmental gradients in space and time.
\end{enumerate}

@@

Perhaps the greatest challenge for ecology in the 21st century is to
forecast the impacts of environmental change (Clark et al. 2001, Petchey
et al. 2015). Forecasts require sophisticated modeling approaches that
fully account for uncertainty and variability in both ecological process
and model parameters (Luo et al. 2011, but see Perretti et al. 2013 for
an argument against modeling the ecological process). The increasing
statistical sophistication of population models (Rees and Ellner 2009)
makes them promising tools for predicting the impacts of environmental
change on species persistence and abundance. But reconciling the scales
at which population models are parameterized and the scales at which
environmental changes play out remains a challenge (Clark et al. 2010,
2012, Freckleton et al. 2011, Queenborough et al. 2011). The problem is
that most population models are built using data from a single study
site because collecting those data, which involves tracking the fates of
individuals plants, is so difficult. The resulting models cannot be
applied to the landscape and regional scales relevant to decision-making
without information about how the fitted parameters respond to spatial
variation in biotic and abiotic drivers (S{æ}ther et al. 2007). The
limited spatial extent of individual-level demographic datasets
constrains our ability to use population models to address applied
questions about the consequences of climate change.

The inability of most population models to address landscape-scale
problems may explain why land managers and conservation planners have
embraced species distribution models (SDMs) (see Guisan and Thuiller
2005 for a review). SDMs typically rely on easy-to-collect
presence/absence data (but see Clark et al. 2014 for new methods) and
remotely-sensed environmental covariates that allow researchers to model
large spatial extents (e.g., Maiorano et al. 2013). Thus, it is
relatively straightforward to parameterize and project SDMs over
landscapes and regions. However, the limitations of SDMs are well known
(Pearson and Dawson 2003, Elith and Leathwick 2009, Ara{ú}jo and
Peterson 2012). Ideally, researchers would provide managers with
landscape-scale population models, combining the extent of SDMs with
information about dynamics and species abundances (Schurr et al. 2012,
Merow et al. 2014).

Aggregate measures of population status, rather than individual
performance, offer an intriguing alternative for modeling populations
(Clark and Bj{ø}rnstad 2004, Freckleton et al. 2011). Population-level
data cannot provide inference about demographic mechanisms, but might be
sufficient for modeling future population states, especially since such
data are feasible to collect across broad spatial extents (e.g.,
Queenborough et al. 2011). The choice between individual and
population-level data involves a difficult trade-off: while
individual-level data leads to more mechanistic models, population-level
data leads to models that can be applied over greater spatial and
temporal extents. An open question is how much forecasting skill is lost
when we build models based on population rather than individual-level
data.

@@

All of our analyses (model fitting and simulating) were conducted in R
(R Core Development Team 2013). We used the `No-U-Turn' MCMC sampler in
Stan (Stan Development Team 2014a) to estimate the posterior
distributions of model parameters using the package `rstan' (Stan
Development Team 2014b). We obtained posterior distributions for all
model parameters from three parallel MCMC chains run for 1,000
iterations after discarding an initial 1,000 iterations. Such short MCMC
chains may surprise readers more familiar with other MCMC samplers
(i.e.~JAGS or WinBUGS), but the Stan sampler is exceptionally efficient,
which reduces the number of iterations needed to achieve convergence. We
assessed convergence visually and made sure scale reduction factors for
all parameters were less than 1.01. For the purposes of including
parameter uncertainty in our population models, we saved the final 1,000
iterations from each of the three MCMC chains to be used as randomly
drawn values during population simulation. This step alleviates the need
to reduce model parameters by model selection since sampling from the
full parameter space in the MCMC ensures that if a parameter broadly
overlaps zero, on average the effect in the population models will also
be near zero. We report the posterior mean, standard deviation, and 95\%
Bayesian Credible Intervals for every parameter of each model for each
species in Appendix B.

@@

Our interpretation is that the QBM is ``missing'' climate signals
associated with at least one vital rate for each species. This leads to
inaccurate and imprecise forecasts because the QBM statistical model
struggles to explain variation due to climate variables that have
positive and negative impacts on different vital rates. When this is the
case, as it is for all our species to varying degrees (Fig. 3),
forecasts from models based on population-level data will fail. Our
result is consistent with related work on the importance of
individual-level data to forecast population responses to exogenous
drivers (Clark et al. 2011a, 2011b, 2012, Galv{á}n et al. 2014).

@@

Adler, P. B., S. P. Ellner, and J. M. Levine. 2010. Coexistence of
perennial plants: An embarrassment of niches.

@@

Ara{ú}jo, M. B., and A. T. Peterson. 2012. Uses and misuses of
bioclimatic envelope modeling. Ecology 93:1527--1539.

@@

Clark, J. S., and O. N. Bj{ø}rnstad. 2004. Population time series:
Process variability, observation errors, missing values, lags, and
hidden states. Ecology 85:3140--3150.

@@

Clark, J. S., D. Bell, C. Chu, B. Courbaud, M. Dietze, M. Hersh, J.
HilleRisLambers, I. Ib{á}{ñ}ez, S. LaDeau, S. McMahon, J. Metcalf, J.
Mohan, E. Moran, L. Pangle, S. Pearson, C. Salk, Z. Shen, D. Valle, and
P. Wyckoff. 2010. High-dimensional coexistence based on individual
variation: a synthesis of evidence. Ecological Monographs 80:569--608.

@@

Galv{á}n, J. D., J. J. Camarero, and E. Guti{é}rrez. 2014. Seeing the
trees for the forest: Drivers of individual growth responses to climate
in Pinus uncinata mountain forests. Journal of Ecology 102:1244--1257.

@@

Petchey, O. L., M. Pontarp, T. M. Massie, S. K{é}fi, A. Ozgul, M.
Weilenmann, G. M. Palamara, F. Altermatt, B. Matthews, J. M. Levine, D.
Z. Childs, B. J. McGill, M. E. Schaepman, B. Schmid, P. Spaak, A. P.
Beckerman, F. Pennekamp, and I. S. Pearse. 2015. The ecological forecast
horizon, and examples of its uses and determinants. Ecology Letters
18:597--611.

@@

R Core Development Team. 2013. R: A language and environment for
statistical computing.

@@

Schurr, F. M., J. Pagel, J. S. Cabral, J. Groeneveld, O. Bykova, R. B.
O'Hara, F. Hartig, W. D. Kissling, H. P. Linder, G. F. Midgley, B.
Schr{ö}der, A. Singer, and N. E. Zimmermann. 2012. How to understand
species' niches and range dynamics: A demographic research agenda for
biogeography. Journal of Biogeography 39:2146--2162.

@@

S{æ}ther, B. E., S. Engen, V. Gr{ø}tan, W. Fiedler, E. Matthysen, M. E.
Visser, J. Wright, A. P. M{ø}ller, F. Adriaensen, H. {Van Balen}, D.
Balmer, M. C. Mainwaring, R. H. McCleery, M. Pampus, and W. Winkel.
2007. The extended Moran effect and large-scale synchronous fluctuations
in the size of great tit and blue tit populations. Journal of Animal
Ecology 76:315--325.

>>>>

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Rapid climate change has generated growing interest in forecasts of
  future population trajectories. Traditional population models,
  typically built using detailed demographic observations from one study
  site, can address climate change impacts at one location, but are
  difficult to scale up to the landscape and regional scales relevant to
  management decisions.
\item
  An alternative is to build models using population-level data that is
  much easier to collect over broad spatial scales than individual-level
  data. However, such models ignore the fact that climate drives
  population growth through its influence on individual performance.
\item
  We test the consequences of aggregating individual responses when
  forecasting climate change impacts on four perennial grass species. We
  parameterized two population models for each species, one based on
  individual-level data (survival, growth and recruitment) and one on
  population-level data (percent cover), and compared their accuracy,
  precision, and sensitivity to climate variables. For both model types
  we used Bayesian ridge regression to identify the optimal predictive
  model in terms of climate covariate strengths.
\item
  The individual-level model was more accurate and precise than the
  aggregated model when predicting out-of-sample observations. When
  comparing climate effects from both models, the population-level model
  missed important climate effects from at least one vital rate for each
  species. Increasing the sample size at the population-level would not
  necessarily reduce forecast uncertainty; the way to reduce uncertainty
  is to capture unique climate dependence of individual vital rates.
\item
  \emph{Synthesis} Our analysis indicates that there is no shortcut to
  forecasting climate change impacts on plant populations --- detailed
  demographic data is essential. Despite the superiority of the
  individual-level model, the forecasts it generated still were too
  uncertain to be useful for decision-makers. We need new methods to
  collect demographic data efficiently across environmental gradients in
  space and time.
\end{enumerate}

@@

Perhaps the greatest challenge for ecology in the 21st century is to
forecast the impacts of environmental change (Clark et al. 2001, Petchey
et al. 2015). Forecasts require sophisticated modeling approaches that
fully account for uncertainty and variability in both ecological process
and model parameters (Luo et al. 2011, but see Perretti et al. 2013 for
an argument against modeling the ecological process). The increasing
statistical sophistication of population models (Rees and Ellner 2009)
makes them promising tools for predicting the impacts of environmental
change on species persistence and abundance. But reconciling the scales
at which population models are parameterized and the scales at which
environmental changes play out remains a challenge (Clark et al. 2010,
2012, Freckleton et al. 2011, Queenborough et al. 2011). The problem is
that most population models are built using data from a single study
site because collecting those data, which involves tracking the fates of
individuals plants, is so difficult. The resulting models cannot be
applied to the landscape and regional scales relevant to decision-making
without information about how the fitted parameters respond to spatial
variation in biotic and abiotic drivers (Sæther et al. 2007). The
limited spatial extent of individual-level demographic datasets
constrains our ability to use population models to address applied
questions about the consequences of climate change.

The inability of most population models to address landscape-scale
problems may explain why land managers and conservation planners have
embraced species distribution models (SDMs) (see Guisan and Thuiller
2005 for a review). SDMs typically rely on easy-to-collect
presence/absence data (but see Clark et al. 2014 for new methods) and
remotely-sensed environmental covariates that allow researchers to model
large spatial extents (e.g., Maiorano et al. 2013). Thus, it is
relatively straightforward to parameterize and project SDMs over
landscapes and regions. However, the limitations of SDMs are well known
(Pearson and Dawson 2003, Elith and Leathwick 2009, Araújo and Peterson
2012). Ideally, researchers would provide managers with landscape-scale
population models, combining the extent of SDMs with information about
dynamics and species abundances (Schurr et al. 2012, Merow et al. 2014).

Aggregate measures of population status, rather than individual
performance, offer an intriguing alternative for modeling populations
(Clark and Bjørnstad 2004, Freckleton et al. 2011). Population-level
data cannot provide inference about demographic mechanisms, but might be
sufficient for modeling future population states, especially since such
data are feasible to collect across broad spatial extents (e.g.,
Queenborough et al. 2011). The choice between individual and
population-level data involves a difficult trade-off: while
individual-level data leads to more mechanistic models, population-level
data leads to models that can be applied over greater spatial and
temporal extents. An open question is how much forecasting skill is lost
when we build models based on population rather than individual-level
data.

@@

All of our analyses (model fitting and simulating) were conducted in R
(R Core Team 2013). We used the `No-U-Turn' MCMC sampler in Stan (Stan
Development Team 2014a) to estimate the posterior distributions of model
parameters using the package `rstan' (Stan Development Team 2014b). We
obtained posterior distributions for all model parameters from three
parallel MCMC chains run for 1,000 iterations after discarding an
initial 1,000 iterations. Such short MCMC chains may surprise readers
more familiar with other MCMC samplers (i.e.~JAGS or WinBUGS), but the
Stan sampler is exceptionally efficient, which reduces the number of
iterations needed to achieve convergence. We assessed convergence
visually and made sure scale reduction factors for all parameters were
less than 1.01. For the purposes of including parameter uncertainty in
our population models, we saved the final 1,000 iterations from each of
the three MCMC chains to be used as randomly drawn values during
population simulation. This step alleviates the need to reduce model
parameters by model selection since sampling from the full parameter
space in the MCMC ensures that if a parameter broadly overlaps zero, on
average the effect in the population models will also be near zero. We
report the posterior mean, standard deviation, and 95\% Bayesian
Credible Intervals for every parameter of each model for each species in
Appendix B.

@@

Our interpretation is that the QBM is ``missing'' climate signals
associated with at least one vital rate for each species. This leads to
inaccurate and imprecise forecasts because the QBM statistical model
struggles to explain variation due to climate variables that have
positive and negative impacts on different vital rates. When this is the
case, as it is for all our species to varying degrees (Fig. 3),
forecasts from models based on population-level data will fail. Our
result is consistent with related work on the importance of
individual-level data to forecast population responses to exogenous
drivers (Clark et al. 2011a, 2011b, 2012, Galván et al. 2014).

@@

Adler, P. B., S. P. Ellner, and J. M. Levine. 2010. Coexistence of
perennial plants: An embarrassment of niches. Ecology Letters
13:1019--1029.

@@

Araújo, M. B., and A. T. Peterson. 2012. Uses and misuses of bioclimatic
envelope modeling. Ecology 93:1527--1539.

@@

Clark, J. S., and O. N. Bjørnstad. 2004. Population time series: Process
variability, observation errors, missing values, lags, and hidden
states. Ecology 85:3140--3150.

@@

Clark, J. S., D. Bell, C. Chu, B. Courbaud, M. Dietze, M. Hersh, J.
HilleRisLambers, I. Ibáñez, S. LaDeau, S. McMahon, J. Metcalf, J. Mohan,
E. Moran, L. Pangle, S. Pearson, C. Salk, Z. Shen, D. Valle, and P.
Wyckoff. 2010. High-dimensional coexistence based on individual
variation: a synthesis of evidence. Ecological Monographs 80:569--608.

@@

Galván, J. D., J. J. Camarero, and E. Gutiérrez. 2014. Seeing the trees
for the forest: Drivers of individual growth responses to climate in
Pinus uncinata mountain forests. Journal of Ecology 102:1244--1257.

@@

Petchey, O. L., M. Pontarp, T. M. Massie, S. Kéfi, A. Ozgul, M.
Weilenmann, G. M. Palamara, F. Altermatt, B. Matthews, J. M. Levine, D.
Z. Childs, B. J. McGill, M. E. Schaepman, B. Schmid, P. Spaak, A. P.
Beckerman, F. Pennekamp, and I. S. Pearse. 2015. The ecological forecast
horizon, and examples of its uses and determinants. Ecology Letters
18:597--611.

@@

R Core Team. 2013. R: A language and environment for statistical
computing.

@@

Schurr, F. M., J. Pagel, J. S. Cabral, J. Groeneveld, O. Bykova, R. B.
O'Hara, F. Hartig, W. D. Kissling, H. P. Linder, G. F. Midgley, B.
Schröder, A. Singer, and N. E. Zimmermann. 2012. How to understand
species' niches and range dynamics: A demographic research agenda for
biogeography. Journal of Biogeography 39:2146--2162.

@@

Sæther, B. E., S. Engen, V. Grøtan, W. Fiedler, E. Matthysen, M. E.
Visser, J. Wright, A. P. Møller, F. Adriaensen, H. {Van Balen}, D.
Balmer, M. C. Mainwaring, R. H. McCleery, M. Pampus, and W. Winkel.
2007. The extended Moran effect and large-scale synchronous fluctuations
in the size of great tit and blue tit populations. Journal of Animal
Ecology 76:315--325.
